---
title: "TikTok Video View Summary"
format: pdf
---

# Setup & Data Import

```{r}
library(dplyr)
library(tidyverse)

views <- read.csv("data/raw/video_view.csv")
```

# Data Inspection

```{r}
count_rows <- nrow(views)
print(count_rows)

str(views)
head(views)
summary(views)
colSums(is.na(views))

names(views)
glimpse(views)
```

# General Summary

```{r}
overview <- views %>%
  summarize(
    videos = n(),
    creators = n_distinct(creator_id),
    avg_length = mean(video_length_sec),
    avg_impressions = mean(impressions_n),
    avg_watched = mean(watched_n),
    avg_watch_rate = mean(watch_rate, na.rm=TRUE),
    avg_watch_share = mean(avg_watch_share, na.rm=TRUE),
    avg_watch_seconds = mean(avg_watch_seconds_when_watched, na.rm=TRUE) 
  )

print(overview)
```

```{r}
views$watched_pct <- views$watched_n / views$impressions_n * 100

hist(views$video_length_sec)

goodvideos <- subset(views, watch_rate > 0.8)
nrow(goodvideos)
```

# Top Videos

### Top Video by Impressions
```{r}
topvideo <- views[views$impressions_n == max(views$impressions_n), ]
print(topvideo)
```

### Top Videos by Watch Rate
```{r}
top_watch <- views %>%
  select(video_id, creator_id, watch_rate, impressions_n) %>%
  filter(impressions_n >= 60) %>%
  arrange(desc(watch_rate))

head(top_watch)
```

# Creator Performance

```{r}
creator_summary <- views %>%
  group_by(creator_id) %>%
  summarise(
    videos_n = n(),
    impressions_total = sum(impressions_n, na.rm = TRUE),
    watch_rate_avg = mean(watch_rate, na.rm = TRUE)
  ) %>%
  arrange(desc(impressions_total))

head(creator_summary)
```

```{r}
creator_engagement <- views %>%
  mutate(engagement = watched_n / impressions_n) %>%
  group_by(creator_id) %>%
  summarise(mean_engagement = mean(engagement, na.rm = TRUE)) %>%
  arrange(desc(mean_engagement))

head(creator_engagement)
```
# Regression Analysis

This analysis investigates the relationship between `score_satiation_penalty` and `feed_rank`.

```{r}
impressions <- read_csv("data/raw/impressions_clean.csv")

model1 <- lm(
  feed_rank ~ score_satiation_penalty,
  data = impressions
)

summary(model1)
```

The first model shows a significant positive relationship between `score_satiation_penalty` and `feed_rank`. A one-unit increase in `score_satiation_penalty` is associated with an increase of approximately 4.09 in `feed_rank`. The model explains approximately 5.8% of the variation in `feed_rank`.

## Relationship Plot

```{r}
ggplot(
  impressions,
  aes(x = score_satiation_penalty, y = feed_rank)
) +
  geom_point(alpha = 0.2) +
  labs(
    title = "Satiation penalty and feed rank",
    x = "Satiation penalty",
    y = "Feed rank"
  ) +
  theme_minimal()
```

The plot shows the positive relationship between `score_satiation_penalty` and `feed_rank` found in the first regression model.

## Extended Regression Model

```{r}
model2 <- lm(
  feed_rank ~ score_satiation_penalty + score_category_match + score_creator_match,
  data = impressions
)

summary(model2)
```

After adding `score_category_match` and `score_creator_match`, the relationship between `score_satiation_penalty` and `feed_rank` remains similar. The extended model explains approximately 5.9% of the variation in `feed_rank`.