---
title: "TikTok Video View Summary"
format: pdf
execute:
  echo: false
  warning: false
---

# Setup & Data Import
This section loads the necessary R libraries and imports the raw TikTok datasets for our analysis.

```{r}
library(dplyr)
library(tidyverse)
library(ggplot2)
library(here)
library(DBI)
library(RSQLite)

if (!dir.exists("data/raw")) {
  dir.create("data/raw", recursive = TRUE)
}

data_url <- paste0("https://raw.githubusercontent.com/hannesdatta/course-dprep/refs/heads/main/material/project/video_view.csv")
download.file(data_url, "data/raw/video_view.csv", mode = "wb")
list.files()

views <- read.csv("data/raw/video_view.csv")

data_url <- "https://raw.githubusercontent.com/hannesdatta/course-dprep/refs/heads/main/material/project/coaching_2_data/users.csv"

download.file(data_url, "data/raw/users.csv", mode = "wb")
list.files()

data_url <- "https://raw.githubusercontent.com/hannesdatta/course-dprep/refs/heads/main/material/project/coaching_2_data/impressions.csv"

download.file(data_url, "data/raw/impressions.csv", mode = "wb")
list.files()

impressions <- read.csv("data/raw/impressions.csv")

data_url <- "https://raw.githubusercontent.com/hannesdatta/course-dprep/refs/heads/main/material/project/coaching_2_data/sessions.csv"

download.file(data_url, "data/raw/sessions.csv", mode = "wb")
list.files()

data_url <- "https://raw.githubusercontent.com/hannesdatta/course-dprep/refs/heads/main/material/project/coaching_2_data/watch_events.csv"

download.file(data_url, "data/raw/watch_events.csv", mode = "wb")
list.files()
```

# Data Inspection
Here we check the size and structure of our data to make sure everything looks right.

```{r}
count_rows <- nrow(views)
print(count_rows)

str(views)
head(views)
summary(views)
colSums(is.na(views))

names(views)
glimpse(views)
```

# General Summary
This section gives an overview of the main numbers, such as average video length, impressions, and watch rates.

```{r}
overview <- views %>%
  summarize(
    videos = n(),
    creators = n_distinct(creator_id),
    avg_length = mean(video_length_sec),
    avg_impressions = mean(impressions_n),
    avg_watched = mean(watched_n),
    avg_watch_rate = mean(watch_rate, na.rm=TRUE),
    avg_watch_share = mean(avg_watch_share, na.rm=TRUE),
    avg_watch_seconds = mean(avg_watch_seconds_when_watched, na.rm=TRUE) 
  )

print(overview)
```

The histogram shows how video lengths are distributed. We also found r nrow(goodvideos) videos with a watch rate higher than 80%.

```{r}
views$watched_pct <- views$watched_n / views$impressions_n * 100

hist(views$video_length_sec)

goodvideos <- subset(views, watch_rate > 0.8)
nrow(goodvideos)
```


# Top Videos
This video got the most impressions in the whole dataset.

### Top Video by Impressions
```{r}
topvideo <- views[views$impressions_n == max(views$impressions_n), ]
print(topvideo)
```

To filter out noise, we look at videos with at least 60 impressions and rank them by their watch rate.

### Top Videos by Watch Rate
```{r}
top_watch <- views %>%
  select(video_id, creator_id, watch_rate, impressions_n) %>%
  filter(impressions_n >= 60) %>%
  arrange(desc(watch_rate))

head(top_watch)
```

# Creator Performance
In this section, we evaluate creator productivity by looking at total impressions generated per creator and their average audience engagement.

```{r}
creator_summary <- views %>%
  group_by(creator_id) %>%
  summarise(
    videos_n = n(),
    impressions_total = sum(impressions_n, na.rm = TRUE),
    watch_rate_avg = mean(watch_rate, na.rm = TRUE)
  ) %>%
  arrange(desc(impressions_total))

head(creator_summary)
```

```{r}
creator_engagement <- views %>%
  mutate(engagement = watched_n / impressions_n) %>%
  group_by(creator_id) %>%
  summarise(mean_engagement = mean(engagement, na.rm = TRUE)) %>%
  arrange(desc(mean_engagement))

head(creator_engagement)
```

# Basic data Inspection
This section inspects and analyses separate datasets, namely the impressions, sessions, users and watch events.

## Impressions analysis

```{r}
impressions <- read.csv("data/raw/impressions.csv")

impressions_clean <- impressions %>%
  distinct(impression_id, .keep_all = TRUE)

write_csv(impressions_clean, "data/raw/impressions_clean.csv")

impressions <- read_csv("data/raw/impressions_clean.csv")

ggplot(impressions, aes(x = source_bucket)) +
  geom_bar() +
  labs(
    title = "Mix of feed sources",
    x = "Feed source",
    y = "Count"
  ) +
  theme_minimal()

ggplot(impressions, aes(x = score_category_match, y = score_total)) +
  geom_point() +
  labs(
    title = "Category match and total ranking score",
    x = "Category match score",
    y = "Total ranking score"
  ) +
  theme_minimal()

ggplot(impressions, aes(x = score_creator_match, y = score_total)) +
  geom_point() +
  labs(
    title = "Creator match and total ranking score",
    x = "Creator match score",
    y = "Total ranking score"
  ) +
  theme_minimal()

ggplot(impressions, aes(x = score_satiation_penalty)) +
  geom_histogram(bins = 30) +
  labs(
    title = "Distribution of satiation penalty",
    x = "Satiation penalty",
    y = "Count"
  ) +
  theme_minimal()

ggplot(impressions, aes(x = score_total)) +
  geom_histogram(bins = 30) +
  labs(
    title = "Distribution of total ranking scores",
    x = "Total ranking score",
    y = "Count"
  ) +
  theme_minimal()
```

## Session analysis

```{r}
# read data
sessions <- read.csv("data/raw/sessions.csv")

# clean data
sessions_clean <- sessions %>%
  filter(!is.na(session_id)) %>%
  distinct() %>%
  select(-matches("name|first_name|last_name", ignore.case = TRUE))

# save output
write_csv(sessions_clean, "data/raw/sessions_cleaned.csv")
print("done cleaning")

# read cleaned dataset
sessions <- read_csv("data/raw/sessions_cleaned.csv")

# Only drop for ID for instance
sessions_clean <- sessions %>%
  drop_na(session_id) %>%
  distinct()

# Rank sessions from longest to shortest
rank_sessions <- sessions %>%
  mutate(
    duration_rank = rank(-session_duration_sec)
  )

p4 <- rank_sessions %>%
  filter(duration_rank <= 15) %>%
  ggplot(aes(x = duration_rank, y = session_duration_sec)) +
  geom_col() +
  labs(
    title = "Top 15 Longest Sessions",
    x = "Rank",
    y = "Session duration (seconds)"
  ) +
  theme_minimal()

# plot 1: distribution of session duration (in minutes)
p1 <- ggplot(sessions, aes(x = session_duration_sec / 60)) +
  geom_histogram(bins = 30, fill = "steelblue", color = "white") +
  labs(
    title = "Distribution of TikTok Session Duration",
    subtitle = "Calculated in minutes per user session",
    x = "Session duration (minutes)",
    y = "Number of sessions"
  ) +
  theme_minimal()

# prepare groups with ordered factor
tier_order <- c("Low engagement", "Medium engagement", "High engagement")

sessions_plot <- sessions %>%
  mutate(
    engagement_tier = case_when(
      videos_viewed < 8 ~ "Low engagement",
      videos_viewed <= 15 ~ "Medium engagement",
      TRUE ~ "High engagement"
    ),
    engagement_tier = factor(engagement_tier, levels = tier_order)
  )

# plot 2: duration vs videos viewed with ordered tiers
p2 <- ggplot(sessions_plot, aes(x = session_duration_sec / 60, y = videos_viewed, color = engagement_tier)) +
  geom_point(alpha = 0.6, size = 1.8) +
  scale_color_discrete(breaks = tier_order) +
  labs(
    title = "Session Duration vs Videos Viewed",
    subtitle = "User engagement tiers across session length",
    x = "Session duration (minutes)",
    y = "Videos viewed",
    color = "Engagement"
  ) +
  theme_minimal()

# plot 3: group comparison by user engagement
sessions_grouped <- sessions_plot %>%
  group_by(engagement_tier) %>%
  summarise(avg_duration_min = mean(session_duration_sec / 60, na.rm = TRUE)) %>%
  mutate(engagement_tier = factor(engagement_tier, levels = tier_order))

p3 <- ggplot(sessions_grouped, aes(x = engagement_tier, y = avg_duration_min, fill = engagement_tier)) +
  geom_col(show.legend = FALSE) +
  scale_x_discrete(limits = tier_order) +
  scale_fill_brewer(palette = "Blues") +
  labs(
    title = "Average Session Duration by Engagement Level",
    subtitle = "Comparing session length across different user tiers",
    x = "Engagement level",
    y = "Average duration (minutes)"
  ) +
  theme_minimal()

p1
p2
p3
p4
```

## Users analysis

```{r}
users <- read.csv("data/raw/users.csv")

users_clean <- users %>%
  select(-user_name, -user_handle)
write.csv(users_clean, "data/raw/users_clean.csv", row.names = FALSE)

# Load the data
users_clean <- read.csv("data/raw/users_clean.csv")

## Create plot 1
# Make plot_1
plot_1 <- ggplot(data = users_clean, aes(x = baseline_login, y = base_videos_watched_mean)) +
  geom_point(color = "blue", alpha = 0.7, size = 3) +
  labs(
    title = "Login frequency vs videos watched",
    x = "Login frequency",
    y = "Mean videos watched"
  ) +
theme_minimal() +
  theme(
    plot.title = element_text(face = "bold", size = 18, color = "white"),
    axis.title = element_text(face = "bold", color = "white"),
    axis.text = element_text(color = "white"),
    panel.grid.major = element_line(color = "#444444"),
    panel.grid.minor = element_blank(),
    plot.background = element_rect(fill = "black", color = NA),
    panel.background = element_rect(fill = "black", color = NA)
  )

## Create plot 2
# Put all content categories in one dataframe
df <- data.frame(
  category = c(
    "Comedy", "Dance", "BeautyFashion", "Food",
    "FitnessSports", "Gaming", "DIYHome", "Travel",
    "Education", "Pets"
  ),
  score = c(
    mean(users_clean$pref_Comedy, na.rm = TRUE),
    mean(users_clean$pref_Dance, na.rm = TRUE),
    mean(users_clean$pref_BeautyFashion, na.rm = TRUE),
    mean(users_clean$pref_Food, na.rm = TRUE),
    mean(users_clean$pref_FitnessSports, na.rm = TRUE),
    mean(users_clean$pref_Gaming, na.rm = TRUE),
    mean(users_clean$pref_DIYHome, na.rm = TRUE),
    mean(users_clean$pref_Travel, na.rm = TRUE),
    mean(users_clean$pref_Education, na.rm = TRUE),
    mean(users_clean$pref_Pets, na.rm = TRUE)
  )
)
# Make plot_2
plot_2 <- ggplot(df, aes(x = category, y = score)) +
  geom_col(fill = "purple", alpha = 0.85) +
  labs(
    title = "Mean preference score across 10 content categories",
    x = "Content category",
    y = "Mean preference score"
  ) +
  theme_light() +
  theme(plot.title = element_text(face = "bold", size = 20),
    axis.text.x = element_text(angle = 45, hjust = 1, size = 12),
    panel.grid.major = element_line(color = "white")
  )

## Create plot 3
# Make plot_3
plot_3 <- ggplot(users_clean, aes(x = satiation_decay)) +
  geom_density(fill = "orange", alpha = 0.6) +
  labs(
    title = "Density of satiation decay",
    x = "Satiation decay",
    y = "Density"
  ) +
  theme_bw() +
  theme(
    plot.title = element_text(face = "italic", size = 18),
    axis.title = element_text(size = 15),
    panel.border = element_rect(color = "orange", fill = NA, size = 1),
    panel.grid.minor = element_blank()
  )

## Create plot 4
# Make plot_4
plot_4 <- ggplot(users_clean, aes(x = need_interaction)) +
  geom_histogram(bins = 30, fill = "green", color = "white", alpha = 0.9) +
  labs(
    title = "Distribution of need for interaction",
    x = "Need for interaction",
    y = "Count"
  ) +
  theme_dark() +
  theme(
    plot.title = element_text(face = "bold", size = 20, color = "darkgreen"),
    axis.title = element_text(face = "bold", size = 18, color = "darkgreen"),
    axis.text = element_text(face = "bold", size = 18, color = "darkgreen")
  )

plot_1
plot_2
plot_3
plot_4
```

## Watch events analysis

```{r}
# 1. Download data
watch_events <- read.csv("data/raw/watch_events.csv")

# 2. Data Cleaning & Handling NA / Missing Values
watch_clean <- watch_events %>%
  select(-started_at_raw) %>%
  filter(!is.na(session_id), !is.na(video_id), !is.na(started_at)) %>%
  mutate(
    watch_seconds = replace_na(watch_seconds, 0),
    started_at_ts = as.POSIXct(started_at, format = "%Y-%m-%dT%H:%M:%SZ", tz = "UTC"),
    started_day = as.Date(started_at_ts),
    action_type = case_when(
      action == "watch_full" ~ "Complete Watch",
      action %in% c("like", "share", "comment") ~ "Active Engagement",
      is.na(action) ~ "Unknown Action",
      TRUE ~ "Browse / Other"
    )
  )

# Save cleaned dataset
write_csv(watch_clean, "data/raw/watch_clean.csv")

# Graph 1: Action distribution
action_count <- watch_clean %>%
  count(action)

action_counts_plot <- ggplot(action_count, aes(x = action, y = n)) +
  geom_col(fill = "steelblue") +
  labs(
    title = "Distribution of how often each action occurs",
    x = "Action",
    y = "Count"
  ) +
  theme_minimal()

# Graph 2: Mean watch time by action
action_group <- watch_clean %>%
  group_by(action) %>%
  summarise(
    mean_seconds = mean(watch_seconds, na.rm = TRUE)
  )

group_action_plot <- ggplot(action_group, aes(x = action, y = mean_seconds)) +
  geom_col(fill = "darkorange") +
  labs(
    title = "Mean watch time by action",
    x = "Action",
    y = "Mean watch time (seconds)"
  ) +
  theme_minimal()

# Graph 3: Number of watches per day
time_watch <- watch_clean %>%
  count(started_day)

time_watches_plot <- ggplot(time_watch, aes(x = started_day, y = n)) +
  geom_line(color = "forestgreen", linewidth = 1) +
  labs(
    title = "Number of watches per day",
    x = "Date",
    y = "Count"
  ) +
  theme_minimal()

# Graph 4: Top sessions with rank()
session_ranking <- watch_clean %>%
  filter(action == "watch_full") %>%
  count(session_id, name = "full_watches") %>%
  mutate(session_rank = rank(-full_watches, ties.method = "min")) %>%
  arrange(session_rank) %>%
  slice_head(n = 15)

plot_4 <- ggplot(session_ranking, aes(x = reorder(factor(session_id), -full_watches), y = full_watches)) +
  geom_col(fill = "purple") +
  labs(
    title = "Top 15 sessions with the most full watches",
    x = "Session ID",
    y = "Number of full watches"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Graph 5: Top 10 most watched videos
video_count <- watch_clean %>%
  filter(action == "watch_full") %>%
  count(video_id, name = "full_watches") %>%
  arrange(desc(full_watches)) %>%
  slice_head(n = 10)

plot_5 <- ggplot(video_count, aes(x = reorder(factor(video_id), -full_watches), y = full_watches)) +
  geom_col(fill = "coral") +
  labs(
    title = "Top 10 most watched videos",
    x = "Video ID",
    y = "Number of full watches"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

action_counts_plot
group_action_plot
time_watches_plot
plot_4
plot_5
```

# Regression Analysis
This analysis investigates the relationship between `score_satiation_penalty` and `feed_rank`.

```{r}
impressions <- read_csv("data/raw/impressions_clean.csv", show_col_types = FALSE)

model1 <- lm(
  feed_rank ~ score_satiation_penalty,
  data = impressions
)

summary(model1)
```
The first model shows a significant positive relationship between `score_satiation_penalty` and `feed_rank`. A one-unit increase in `score_satiation_penalty` is associated with an increase of approximately 4.09 in `feed_rank`. The model explains approximately 5.8% of the variation in `feed_rank`.

## Relationship Plot
The plot above visualizes the positive relationship between score_satiation_penalty and feed_rank found in our initial regression model, complete with a linear trend line

```{r}
ggplot(
  impressions,
  aes(x = score_satiation_penalty, y = feed_rank)
) +
  geom_point(alpha = 0.2) +
  geom_smooth(method = "lm", color = "blue", se = TRUE) +
  labs(
    title = "Satiation penalty and feed rank",
    x = "Satiation penalty",
    y = "Feed rank"
  ) +
  theme_minimal()
```

The plot shows the positive relationship between `score_satiation_penalty` and `feed_rank` found in the first regression model.

## Extended Regression Model
```{r}
model2 <- lm(
  feed_rank ~ score_satiation_penalty + score_category_match + score_creator_match,
  data = impressions
)

summary(model2)
```

After adding `score_category_match` and `score_creator_match`, the relationship between `score_satiation_penalty` and `feed_rank` remains similar. The extended model explains approximately 5.9% of the variation in `feed_rank`.

# SQLlite analysis
Separate tables were created from the datasets from the SQL database.

```{r}
# Download SQLite file from the provided URL
url <- "https://filesender.surf.nl/download.php??token=29803da2-2322-4844-aebf-7e0b95129957&files_ids=38390042"
destination_file <- "data/raw/tiktok_students.sqlite"

# Download the file
download.file(url, destfile = destination_file, mode = "wb")

cat("File downloaded successfully to:", destination_file, "\n")

con <- dbConnect(SQLite(), dbname = "data/raw/tiktok_students.sqlite")
interactions <- dbGetQuery(con, "SELECT * FROM interactions") %>% tibble()
dbDisconnect(con)
print(interactions)

con <- dbConnect(SQLite(), dbname = "data/raw/tiktok_students.sqlite")
videos <- dbGetQuery(con, "SELECT
  video_id,
  video_length_sec,
  COUNT(*) AS videos_amount
FROM videos
GROUP BY creator_id
ORDER BY videos_amount DESC;") %>% tibble()
dbDisconnect(con)
print(videos)
```