---
title: "TikTok Video View Summary"
format: pdf
execute:
  echo: false
  warning: false
---
# Introduction
This report summarizes the analysis of TikTok video view data, providing insights into user engagement, video performance, and creator productivity. The analysis includes data inspection, summary statistics, visualizations, regression analysis, and SQL database queries to extract relevant information.

# Setup & Data Import
This section loads the necessary R libraries and imports the raw TikTok datasets for our analysis. Downloading the files inside the Quarto document helps make the report reproducible.

```{r}
library(dplyr)
library(tidyverse)
library(ggplot2)
library(here)
library(DBI)
library(RSQLite)

if (!dir.exists("data/raw")) {
  dir.create("data/raw", recursive = TRUE)
}

data_url <- paste0("https://raw.githubusercontent.com/hannesdatta/course-dprep/refs/heads/main/material/project/video_view.csv")
download.file(data_url, "data/raw/video_view.csv", mode = "wb")
list.files()

views <- read.csv("data/raw/video_view.csv")

data_url <- "https://raw.githubusercontent.com/hannesdatta/course-dprep/refs/heads/main/material/project/coaching_2_data/users.csv"

download.file(data_url, "data/raw/users.csv", mode = "wb")
list.files()

data_url <- "https://raw.githubusercontent.com/hannesdatta/course-dprep/refs/heads/main/material/project/coaching_2_data/impressions.csv"

download.file(data_url, "data/raw/impressions.csv", mode = "wb")
list.files()

impressions <- read.csv("data/raw/impressions.csv")

data_url <- "https://raw.githubusercontent.com/hannesdatta/course-dprep/refs/heads/main/material/project/coaching_2_data/sessions.csv"

download.file(data_url, "data/raw/sessions.csv", mode = "wb")
list.files()

data_url <- "https://raw.githubusercontent.com/hannesdatta/course-dprep/refs/heads/main/material/project/coaching_2_data/watch_events.csv"

download.file(data_url, "data/raw/watch_events.csv", mode = "wb")
list.files()
```

# Data Inspection
Before starting the analysis, we inspect the video view dataset regarding its size and structure. This helps us understand what one row represents, which variables are available, whether variables have the expected data types, and whether there are missing values. This step is important because later summaries and plots depend on using the correct variables.

```{r}
count_rows <- nrow(views)
print(count_rows)

str(views)
head(views)
summary(views)
colSums(is.na(views))

names(views)
glimpse(views)
```

# General Summary
This section gives an overview of video view dataset, such as average video length, impressions, and watch rates. These metrics give a first impression of the overall scale and performance of the dataset before looking at individual videos or creators.

```{r}
overview <- views %>%
  summarize(
    videos = n(),
    creators = n_distinct(creator_id),
    avg_length = mean(video_length_sec),
    avg_impressions = mean(impressions_n),
    avg_watched = mean(watched_n),
    avg_watch_rate = mean(watch_rate, na.rm=TRUE),
    avg_watch_share = mean(avg_watch_share, na.rm=TRUE),
    avg_watch_seconds = mean(avg_watch_seconds_when_watched, na.rm=TRUE) 
  )

print(overview)
```

The histogram shows how video lengths are distributed. We also found `r nrow(goodvideos)` videos with a watch rate higher than 80%.

```{r}
views$watched_pct <- views$watched_n / views$impressions_n * 100

hist(views$video_length_sec)

goodvideos <- subset(views, watch_rate > 0.8)
nrow(goodvideos)
```

# Top Videos
This section identifies videos with the most impressions.

### Top Video by Impressions
```{r}
topvideo <- views[views$impressions_n == max(views$impressions_n), ]
print(topvideo)
```

To filter out noise, we look at videos with at least 60 impressions and rank them by their watch rate.

### Top Videos by Watch Rate
```{r}
top_watch <- views %>%
  select(video_id, creator_id, watch_rate, impressions_n) %>%
  filter(impressions_n >= 60) %>%
  arrange(desc(watch_rate))

head(top_watch)
```

# Creator Performance
In this section, we evaluate creator productivity by looking at total impressions generated per creator and their average audience engagement.

```{r}
creator_summary <- views %>%
  group_by(creator_id) %>%
  summarise(
    videos_n = n(),
    impressions_total = sum(impressions_n, na.rm = TRUE),
    watch_rate_avg = mean(watch_rate, na.rm = TRUE)
  ) %>%
  arrange(desc(impressions_total))

head(creator_summary)
```

```{r}
creator_engagement <- views %>%
  mutate(engagement = watched_n / impressions_n) %>%
  group_by(creator_id) %>%
  summarise(mean_engagement = mean(engagement, na.rm = TRUE)) %>%
  arrange(desc(mean_engagement))

head(creator_engagement)
```

# Additional Data Inspection
In addition to the video view dataset, we analyze several related datasets: impressions, sessions, users, and watch events. Each dataset captures a different part of the TikTok user experience. The impressions data describe how videos are distributed in the feed, the sessions data describe the duration of sessions, the users data describe user preferences, and the watch events data describe actual viewing behavior.

## Impressions analysis
The impressions analysis gives insight into how videos are distributed in the feed and how ranking scores behave. The feed source plot shows that `preferred_new` is the most common source of impressions. The relationship between category match and total ranking score is clearly positive, indicating that better category matching is associated with higher total ranking scores. In contrast, the creator match score shows a more scattered relationship with the total score. The distribution plots show that satiation penalties are generally low, while total ranking scores follow a distribution around zero.

```{r}
impressions <- read.csv("data/raw/impressions.csv")

impressions_clean <- impressions %>%
  distinct(impression_id, .keep_all = TRUE)

write_csv(impressions_clean, "data/raw/impressions_clean.csv")

impressions <- read_csv("data/raw/impressions_clean.csv")

ggplot(impressions, aes(x = source_bucket)) +
  geom_bar() +
  labs(
    title = "Mix of feed sources",
    x = "Feed source",
    y = "Count"
  ) +
  theme_minimal()

ggplot(impressions, aes(x = score_category_match, y = score_total)) +
  geom_point() +
  labs(
    title = "Category match and total ranking score",
    x = "Category match score",
    y = "Total ranking score"
  ) +
  theme_minimal()

ggplot(impressions, aes(x = score_creator_match, y = score_total)) +
  geom_point() +
  labs(
    title = "Creator match and total ranking score",
    x = "Creator match score",
    y = "Total ranking score"
  ) +
  theme_minimal()

ggplot(impressions, aes(x = score_satiation_penalty)) +
  geom_histogram(bins = 30) +
  labs(
    title = "Distribution of satiation penalty",
    x = "Satiation penalty",
    y = "Count"
  ) +
  theme_minimal()

ggplot(impressions, aes(x = score_total)) +
  geom_histogram(bins = 30) +
  labs(
    title = "Distribution of total ranking scores",
    x = "Total ranking score",
    y = "Count"
  ) +
  theme_minimal()
```

## Session analysis
The session analysis provides insight into how users spend time on the platform. Session durations are right-skewed, meaning that most sessions are short and only a few are relatively long. The relationship between duration and videos viewed is positive, indicating that longer sessions are associated with higher video consumption. The engagement comparison further supports this: high-engagement sessions have the longest average duration, followed by medium- and low-engagement sessions. Finally, the top-session shows a few long sessions, including one clear outlier.

```{r}
# read data
sessions <- read.csv("data/raw/sessions.csv")

# clean data
sessions_clean <- sessions %>%
  filter(!is.na(session_id)) %>%
  distinct() %>%
  select(-matches("name|first_name|last_name", ignore.case = TRUE))

# save output
write_csv(sessions_clean, "data/raw/sessions_cleaned.csv")
print("done cleaning")

# read cleaned dataset
sessions <- read_csv("data/raw/sessions_cleaned.csv")

# Only drop for ID for instance
sessions_clean <- sessions %>%
  drop_na(session_id) %>%
  distinct()

# Rank sessions from longest to shortest
rank_sessions <- sessions %>%
  mutate(
    duration_rank = rank(-session_duration_sec)
  )

p4 <- rank_sessions %>%
  filter(duration_rank <= 15) %>%
  ggplot(aes(x = duration_rank, y = session_duration_sec)) +
  geom_col() +
  labs(
    title = "Top 15 Longest Sessions",
    x = "Rank",
    y = "Session duration (seconds)"
  ) +
  theme_minimal()

# plot 1: distribution of session duration (in minutes)
p1 <- ggplot(sessions, aes(x = session_duration_sec / 60)) +
  geom_histogram(bins = 30, fill = "steelblue", color = "white") +
  labs(
    title = "Distribution of TikTok Session Duration",
    subtitle = "Calculated in minutes per user session",
    x = "Session duration (minutes)",
    y = "Number of sessions"
  ) +
  theme_minimal()

# prepare groups with ordered factor
tier_order <- c("Low engagement", "Medium engagement", "High engagement")

sessions_plot <- sessions %>%
  mutate(
    engagement_tier = case_when(
      videos_viewed < 8 ~ "Low engagement",
      videos_viewed <= 15 ~ "Medium engagement",
      TRUE ~ "High engagement"
    ),
    engagement_tier = factor(engagement_tier, levels = tier_order)
  )

# plot 2: duration vs videos viewed with ordered tiers
p2 <- ggplot(sessions_plot, aes(x = session_duration_sec / 60, y = videos_viewed, color = engagement_tier)) +
  geom_point(alpha = 0.6, size = 1.8) +
  scale_color_discrete(breaks = tier_order) +
  labs(
    title = "Session Duration vs Videos Viewed",
    subtitle = "User engagement tiers across session length",
    x = "Session duration (minutes)",
    y = "Videos viewed",
    color = "Engagement"
  ) +
  theme_minimal()

# plot 3: group comparison by user engagement
sessions_grouped <- sessions_plot %>%
  group_by(engagement_tier) %>%
  summarise(avg_duration_min = mean(session_duration_sec / 60, na.rm = TRUE)) %>%
  mutate(engagement_tier = factor(engagement_tier, levels = tier_order))

p3 <- ggplot(sessions_grouped, aes(x = engagement_tier, y = avg_duration_min, fill = engagement_tier)) +
  geom_col(show.legend = FALSE) +
  scale_x_discrete(limits = tier_order) +
  scale_fill_brewer(palette = "Blues") +
  labs(
    title = "Average Session Duration by Engagement Level",
    subtitle = "Comparing session length across different user tiers",
    x = "Engagement level",
    y = "Average duration (minutes)"
  ) +
  theme_minimal()

p1
p2
p3
p4
```

## Users analysis
The user plots show that user behavior and preferences are heterogeneous. The first plot shows that login frequency does not appear to have a strong linear relationship with the average number of videos watched, as the observations are widely spread out. Average content preferences differ by category, with some categories receiving positive mean scores and others negative mean scores. The distributions of satiation decay and need for interaction are relatively flat, indicating that these user characteristics are spread evenly across the sample rather than being concentrated around one specific value.

```{r}
users <- read.csv("data/raw/users.csv")

users_clean <- users %>%
  select(-user_name, -user_handle)
write.csv(users_clean, "data/raw/users_clean.csv", row.names = FALSE)

# Load the data
users_clean <- read.csv("data/raw/users_clean.csv")

## Create plot 1
# Make plot_1
plot_1 <- ggplot(data = users_clean, aes(x = baseline_login, y = base_videos_watched_mean)) +
  geom_point(color = "blue", alpha = 0.7, size = 3) +
  labs(
    title = "Login frequency vs videos watched",
    x = "Login frequency",
    y = "Mean videos watched"
  ) +
theme_minimal() +
  theme(
    plot.title = element_text(face = "bold", size = 18, color = "white"),
    axis.title = element_text(face = "bold", color = "white"),
    axis.text = element_text(color = "white"),
    panel.grid.major = element_line(color = "#444444"),
    panel.grid.minor = element_blank(),
    plot.background = element_rect(fill = "black", color = NA),
    panel.background = element_rect(fill = "black", color = NA)
  )

## Create plot 2
# Put all content categories in one dataframe
df <- data.frame(
  category = c(
    "Comedy", "Dance", "BeautyFashion", "Food",
    "FitnessSports", "Gaming", "DIYHome", "Travel",
    "Education", "Pets"
  ),
  score = c(
    mean(users_clean$pref_Comedy, na.rm = TRUE),
    mean(users_clean$pref_Dance, na.rm = TRUE),
    mean(users_clean$pref_BeautyFashion, na.rm = TRUE),
    mean(users_clean$pref_Food, na.rm = TRUE),
    mean(users_clean$pref_FitnessSports, na.rm = TRUE),
    mean(users_clean$pref_Gaming, na.rm = TRUE),
    mean(users_clean$pref_DIYHome, na.rm = TRUE),
    mean(users_clean$pref_Travel, na.rm = TRUE),
    mean(users_clean$pref_Education, na.rm = TRUE),
    mean(users_clean$pref_Pets, na.rm = TRUE)
  )
)
# Make plot_2
plot_2 <- ggplot(df, aes(x = category, y = score)) +
  geom_col(fill = "purple", alpha = 0.85) +
  labs(
    title = "Mean preference score across 10 content categories",
    x = "Content category",
    y = "Mean preference score"
  ) +
  theme_light() +
  theme(plot.title = element_text(face = "bold", size = 20),
    axis.text.x = element_text(angle = 45, hjust = 1, size = 12),
    panel.grid.major = element_line(color = "white")
  )

## Create plot 3
# Make plot_3
plot_3 <- ggplot(users_clean, aes(x = satiation_decay)) +
  geom_density(fill = "orange", alpha = 0.6) +
  labs(
    title = "Density of satiation decay",
    x = "Satiation decay",
    y = "Density"
  ) +
  theme_bw() +
  theme(
    plot.title = element_text(face = "italic", size = 18),
    axis.title = element_text(size = 15),
    panel.border = element_rect(color = "orange", fill = NA, size = 1),
    panel.grid.minor = element_blank()
  )

## Create plot 4
# Make plot_4
plot_4 <- ggplot(users_clean, aes(x = need_interaction)) +
  geom_histogram(bins = 30, fill = "green", color = "white", alpha = 0.9) +
  labs(
    title = "Distribution of need for interaction",
    x = "Need for interaction",
    y = "Count"
  ) +
  theme_dark() +
  theme(
    plot.title = element_text(face = "bold", size = 20, color = "darkgreen"),
    axis.title = element_text(face = "bold", size = 18, color = "darkgreen"),
    axis.text = element_text(face = "bold", size = 18, color = "darkgreen")
  )

plot_1
plot_2
plot_3
plot_4
```

## Watch events analysis
The watch events plots show variation in viewing behavior. Although action types have similar counts, their watch times differ substantially. Full watches have the highest average watch duration, with the others being fairly shorter. Daily watch counts show fluctuation and are first decreasing before increasing over the observed period. The rankings of sessions and videos further show that full watches are not evenly distributed since a small number of sessions and videos account for the highest numbers of complete views.

```{r}
# 1. Download data
watch_events <- read.csv("data/raw/watch_events.csv")

# 2. Data Cleaning & Handling NA / Missing Values
watch_clean <- watch_events %>%
  select(-started_at_raw) %>%
  filter(!is.na(session_id), !is.na(video_id), !is.na(started_at)) %>%
  mutate(
    watch_seconds = replace_na(watch_seconds, 0),
    started_at_ts = as.POSIXct(started_at, format = "%Y-%m-%dT%H:%M:%SZ", tz = "UTC"),
    started_day = as.Date(started_at_ts),
    action_type = case_when(
      action == "watch_full" ~ "Complete Watch",
      action %in% c("like", "share", "comment") ~ "Active Engagement",
      is.na(action) ~ "Unknown Action",
      TRUE ~ "Browse / Other"
    )
  )

# Save cleaned dataset
write_csv(watch_clean, "data/raw/watch_clean.csv")

# Graph 1: Action distribution
action_count <- watch_clean %>%
  count(action)

action_counts_plot <- ggplot(action_count, aes(x = action, y = n)) +
  geom_col(fill = "steelblue") +
  labs(
    title = "Distribution of how often each action occurs",
    x = "Action",
    y = "Count"
  ) +
  theme_minimal()

# Graph 2: Mean watch time by action
action_group <- watch_clean %>%
  group_by(action) %>%
  summarise(
    mean_seconds = mean(watch_seconds, na.rm = TRUE)
  )

group_action_plot <- ggplot(action_group, aes(x = action, y = mean_seconds)) +
  geom_col(fill = "darkorange") +
  labs(
    title = "Mean watch time by action",
    x = "Action",
    y = "Mean watch time (seconds)"
  ) +
  theme_minimal()

# Graph 3: Number of watches per day
time_watch <- watch_clean %>%
  count(started_day)

time_watches_plot <- ggplot(time_watch, aes(x = started_day, y = n)) +
  geom_line(color = "forestgreen", linewidth = 1) +
  labs(
    title = "Number of watches per day",
    x = "Date",
    y = "Count"
  ) +
  theme_minimal()

# Graph 4: Top sessions with rank()
session_ranking <- watch_clean %>%
  filter(action == "watch_full") %>%
  count(session_id, name = "full_watches") %>%
  mutate(session_rank = rank(-full_watches, ties.method = "min")) %>%
  arrange(session_rank) %>%
  slice_head(n = 15)

plot_4 <- ggplot(session_ranking, aes(x = reorder(factor(session_id), -full_watches), y = full_watches)) +
  geom_col(fill = "purple") +
  labs(
    title = "Top 15 sessions with the most full watches",
    x = "Session ID",
    y = "Number of full watches"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Graph 5: Top 10 most watched videos
video_count <- watch_clean %>%
  filter(action == "watch_full") %>%
  count(video_id, name = "full_watches") %>%
  arrange(desc(full_watches)) %>%
  slice_head(n = 10)

plot_5 <- ggplot(video_count, aes(x = reorder(factor(video_id), -full_watches), y = full_watches)) +
  geom_col(fill = "coral") +
  labs(
    title = "Top 10 most watched videos",
    x = "Video ID",
    y = "Number of full watches"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

action_counts_plot
group_action_plot
time_watches_plot
plot_4
plot_5
```

# Regression Analysis
This analysis investigates the relationship between `score_satiation_penalty` and `feed_rank`.

```{r}
impressions <- read_csv("data/raw/impressions_clean.csv", show_col_types = FALSE)

model1 <- lm(
  feed_rank ~ score_satiation_penalty,
  data = impressions
)

summary(model1)
```

The first model shows a significant positive relationship between `score_satiation_penalty` and `feed_rank`. A one-unit increase in `score_satiation_penalty` is associated with an increase of approximately 4.09 in `feed_rank`. The model explains approximately 5.8% of the variation in `feed_rank`.

## Relationship Plot
The plot above visualizes the positive relationship between score_satiation_penalty and feed_rank found in our initial regression model, complete with a linear trend line

```{r}
ggplot(
  impressions,
  aes(x = score_satiation_penalty, y = feed_rank)
) +
  geom_point(alpha = 0.2) +
  geom_smooth(method = "lm", color = "blue", se = TRUE) +
  labs(
    title = "Satiation penalty and feed rank",
    x = "Satiation penalty",
    y = "Feed rank"
  ) +
  theme_minimal()
```

The plot shows the positive relationship between `score_satiation_penalty` and `feed_rank` found in the first regression model.

## Extended Regression Model
```{r}
model2 <- lm(
  feed_rank ~ score_satiation_penalty + score_category_match + score_creator_match,
  data = impressions
)

summary(model2)
```

After adding `score_category_match` and `score_creator_match`, the relationship between `score_satiation_penalty` and `feed_rank` remains similar. The extended model explains approximately 5.9% of the variation in `feed_rank`.

# SQLlite analysis
This section demonstrates how data can also be retrieved directly from a SQLite database. Instead of only working with CSV files, we connect to the database using `DBI` and `RSQLite` and run SQL queries to extract specific tables or aggregated results. The first query retrieves the full `interactions` table. The second query groups videos by creator and counts how many videos are associated with each creator.

```{r}
# Download SQLite file from the provided URL
url <- "https://filesender.surf.nl/download.php??token=29803da2-2322-4844-aebf-7e0b95129957&files_ids=38390042"
destination_file <- "data/raw/tiktok_students.sqlite"

# Download the file
download.file(url, destfile = destination_file, mode = "wb")

cat("File downloaded successfully to:", destination_file, "\n")

con <- dbConnect(SQLite(), dbname = "data/raw/tiktok_students.sqlite")
interactions <- dbGetQuery(con, "SELECT * FROM interactions") %>% tibble()
dbDisconnect(con)
print(interactions)

con <- dbConnect(SQLite(), dbname = "data/raw/tiktok_students.sqlite")
videos <- dbGetQuery(con, "SELECT
  video_id,
  video_length_sec,
  COUNT(*) AS videos_amount
FROM videos
GROUP BY creator_id
ORDER BY videos_amount DESC;") %>% tibble()
dbDisconnect(con)
print(videos)
```

The SQLite analysis confirms that the database contains detailed interaction-level information. The `interactions` table includes 514,909 records, showing that user behavior is captured at a much more granular level than in the aggregated CSV summaries. The video aggregation also suggests that creator activity is unevenly distributed. Some creators have substantially more videos than others, with the most active creators contributing more than 200 videos. This indicates that creator productivity varies strongly across the platform.

# Conclusion
This report combined several TikTok-related datasets into one PDF analysis. The video view data were used to summarize overall video and creator performance, while the impressions, sessions, users, and watch events datasets provided additional insight into feed exposure, user behavior, and viewing activity.

The analysis shows that creator performance can be evaluated in different ways. Total impressions capture reach, while watch rate and engagement measures capture how users respond to shown videos. The session and watch event analyses further show variation in viewing behavior across users, sessions, actions, and videos. The regression analysis suggests an association between ranking-related scores and feed rank. Overall, the report demonstrates how multiple data sources can be combined into a reproducible Quarto PDF workflow.
